---
title: "Presidential Tweets"
author: "Mahima Kaur"
date: "2023-01-19"
output: html_document
---

## Presidential Tweets

### Obtaining Recent Obama Tweets
Firstly, we need a few R 'packages' designed specifically for playing with Tweets.

```{r}
library(rtweet)
library(lubridate)
library(wordcloud)
library(tidytext)
library(dplyr)
library(ggplot2)
library(tidyr)
library(syuzhet)
library(quantmod)
```

For this class assignment we were given the data of the tweets posted by Obama and trump in recent times. The Obama tweets are thpugh 2pm EST, 1.16.23

#### Loading the files 

```{r}
load("~/Downloads/Obama.Rdata")
load("~/Downloads/Trump.Rdata")
```

#### Exploratory Analysis of Obama's Tweet 

##### How many tweets are there?

```{r}
dim(Obamatweets)
```

##### What are the names of the column? 

```{r}
names(Obamatweets)
```

##### What is the start time of Obama tweets?

```{r}
head(Obamatweets$created_at,1)
```
##### What is the  end time of Obama tweets?

```{r}
tail(Obamatweets$created_at, 1)
```

```{r}
Encoding(Obamatweets$text) <- "UTF-8"
```

#### Now we will create a dataframe including only some important variable for analysis and renaming the headers.

```{r}
Obamatweets <- data.frame(id = Obamatweets$id,
                          text = Obamatweets$text,
                          date = with_tz(Obamatweets$created_at, "US/Pacific"),
                          source = Obamatweets$source,
                          is_quote = Obamatweets$is_quote_status,
                          is_retweet = Obamatweets$retweeted,
                          is_reply = !is.na(Obamatweets$in_reply_to_status_id),
                          favorite_count = Obamatweets$favorite_count,
                          retweet_count = Obamatweets$retweet_count,
                          stringsAsFactors = FALSE)
```

####  After not including some of the variables in the data frame we will again see the shape of the datafram 

```{r}
dim(Obamatweets)
```

####  Now, we want to see what is the source device of the tweets, 

```{r}
table(Obamatweets$source)
```

#### We see that it is a lot of messy data so we will now clean the data so that we can get only the device name. 
#### Reference : https://stackoverflow.com/questions/17227294/removing-html-tags-from-a-string-in-r

```{r}
cleanFun <- function(htmlString) {
return(gsub("<.*?>", "", htmlString))
}
```

```{r}
Obamatweets$source <- cleanFun(Obamatweets$source)
table(Obamatweets$source)
```

#### To clean things more we will remove the words 'Twitter' and 'for' from the source headings 

```{r}
Obamatweets$source <- gsub("Twitter |for", "", Obamatweets$source)
table(Obamatweets$source)
```


Findings : Highest Number of tweets were made from Iphone and after that from Web Client. The least tweets were made from ThunderClap. 

#### Now let's look at the top 10 tweets from two devices 

```{r}
head(Obamatweets$text[Obamatweets$source == "Media Studio"], 10)
```

```{r}
head(Obamatweets$text[Obamatweets$source == " iPhone"], 10)
```
#### Trump's Tweet 

#### Exploratory Analysis of Trumps's Tweet 

##### How many tweets are there?

```{r}
dim(tweets)
```

##### What are the names of the column? 

```{r}
names(tweets)
```

##### Now we will sort the tweets in the reverse order and get the recent 2000 tweets.

```{r}
Trumptweets <- tweets[order(tweets$created_at, decreasing = TRUE), ][1:2000, ]
```


#### Now we will create a dataset with same variable as that of the Obama's tweets dataset. 

```{r}
Trumptweets <- data.frame(id = Trumptweets$user_id,
                          text = Trumptweets$text,
                          date = with_tz(Trumptweets$created_at, "US/Pacific"),
                          source = Trumptweets$source,
                          is_quote = Trumptweets$is_quote,
                          is_retweet = Trumptweets$is_retweet,
                          is_reply = !is.na(Trumptweets$reply_to_status_id),
                          favorite_count = Trumptweets$quoted_favorite_count,
                          retweet_count =Trumptweets$quoted_retweet_count,
                          stringsAsFactors = FALSE)

```


##### Removing any duplicates from the datset 

```{r}
Trumptweets <- Trumptweets[!duplicated(Trumptweets), ]
```

##### Now we will recheck How many tweets are there after removing the duplicates?

```{r}
dim(Trumptweets)
```

Findings : There were 19 duplicates which got removed from the datset. 

##### We will now examine the sources/devices from which the tweets were made. 

```{r}
table(Trumptweets$source)
```

##### Removing twitter and for from the source heading

```{r}
Trumptweets$source <- gsub("Twitter ", "", Trumptweets$source)
Trumptweets$source <- gsub("for ", "", Trumptweets$source)
```

##### We will now examine the sources/devices from which the tweets were made after data cleaning. 

```{r}
table(Trumptweets$source)
```

#### A few tweets from each sources

```{r}
head(Trumptweets$text[Trumptweets$source == "iPhone"])
```

```{r}
head(Trumptweets$text[Trumptweets$source == "Media Studio"])
```

## Data Analysis using Plots


#### For Trump, 1981 Tweets covers about 2020-03-18 to 2020-05-21

```{r}
Trumptweets_by_date <- aggregate(Trumptweets$text ~ date(Trumptweets$date),
                      FUN = length)
```

```{r}
colnames(Trumptweets_by_date) <- c("date", "count")
plot(count ~ date, data = Trumptweets_by_date )
```
#### if we want a line plot instead of a scatterplot 

```{r}
plot(count ~ date, data = Trumptweets_by_date, type = "l", xlab = "Date", ylab = "#Tweets", col = "red", lwd = 2)
```

#### To make the x labels finer 


```{r}
plot(count ~ date, data = Trumptweets_by_date, type = "l", lwd = 2, col = "red", xlab = "Date", ylab = "#Tweets", xaxt = "n")
```
#### Now, we want to have a proper date format on the xlabels 

```{r}
range(Trumptweets$date)
```

#### How many days are there between the range?

```{r}
range(date(Trumptweets$date))[2] - range(date(Trumptweets$date))[1]
```


```{r}
xdates <- seq(from = as.Date("2020-03-18"),
              to = as.Date("2020-05-21"), by = "months")
```

```{r}
xdates
```

#### Now, we want the date format to %m/%d/%y

```{r}
plot(count ~ date, data = Trumptweets_by_date, type = "l", lwd = 2, col = "red", xlab = "Date", ylab = "#Tweets", xaxt = "n")
axis(1, labels=format(xdates,"%m/%d/%y"), at = xdates)
```
Findings : From the plot we can see that the max number of tweets were made in the month of late April 2020. 

#### Now we will see Tewwts Volume by Month 

```{r}
Trumptweets_by_month <- table(month(Trumptweets$date))
```

```{r}
barplot(Trumptweets_by_month, xlab = "Month", ylab = "Number of all TrumpTweets", main = "Tweets Volume per Month")
```





Findings : From the barplot we can conclude that the highest volume of the tweets was the in the Month of April 2020. 


#### Trump Tweets by Hour 

```{r}
Trumptweets_by_hour <- table(hour(Trumptweets$date))
```

```{r}
hrs <- as.numeric(names(Trumptweets_by_hour))
```

```{r}
plot(as.numeric(Trumptweets_by_hour) ~ hrs, type = "h", lwd = 2, col = 2, xlab = "Hour (UTC)", ylab = "Number of all TrumpTweets", main = "Tweet Volume by Hour")
```

#### Converting the time zone from UTC to EST 

```{r}
EST_hrs <- hrs - 5
```

```{r}
EST_hrs <- EST_hrs %% 24
```

```{r}
plot(as.numeric(Trumptweets_by_hour) ~ EST_hrs, type = "h", lwd = 2, col = 2, xlab = "Hour (EST)", ylab = "Number of all TrumpTweets", main = "Tweet Volume by Hour")
```

#### Now, Lets see what is teh daily average of the tweets tweeted by by Trump

```{r}
tweetdays <- range(date(Trumptweets$date))[2] - range(date(Trumptweets$date))[1]
```


```{r}
mean_Trumptweets_by_hour <- Trumptweets_by_hour/as.numeric(tweetdays)
```

```{r}
plot(as.numeric(mean_Trumptweets_by_hour) ~ EST_hrs, type = "h", lwd = 2, col = 2, xlab = "Hour (EST)", ylab = "Mean TrumpTweets", main = "Mean Tweet Volume by Hour")
```


#### What is the mean tweets per day? 
```{r}
round(nrow(Trumptweets)/as.numeric(tweetdays),1)
```

#### Now, Lets look at the Obama's Tweets 

##### Obama's Tweets by day

```{r}
Obamatweets_by_date <- aggregate(Obamatweets$text ~ date(Obamatweets$date),
                            FUN = length)
colnames(Obamatweets_by_date) <- c("date", "count")

plot(count ~ date, data = Obamatweets_by_date, 
     type = "l", lwd = 2, col = 2, xlab = "Date",
     ylab = "# Tweets", xaxt="n")
```

#### What is the range of the Obama's Tweets 

```{r}
range(Obamatweets$date)
```

#### Let's look at the years of the Obama's Twwets 

```{r}
xdates <- seq(from = as.Date("2016-04-08"),
              to = as.Date("2023-01-13"), by = "years")
```

```{r}
xdates
```

#### Let's make a plot with the years as the xasis and see the count per year on the y axis

```{r}
plot(count ~ date, data = Obamatweets_by_date, type = "l", lwd = 2, col = 2, xlab = "Date",
     ylab = "# Tweets", xaxt="n")
axis(1, labels = format(xdates, "%m/%y"), at = xdates )
```
#### How many days are in the dataset?

```{r}
Obamatweetdays <- range(date(Obamatweets$date))[2] - range(date(Obamatweets$date))[1]
```

```{r}
Obamatweetdays
```


#### Lets see on average how many tweets are made per day 

```{r}
round(nrow(Obamatweets)/as.numeric(Obamatweetdays),3)
```

### Trumps'Tweet Popularity 


```{r}
pop_by_day <- aggregate(cbind(favorite_count, retweet_count) ~ as.Date(date), data = Trumptweets, FUN = median)
head(pop_by_day)

colnames(pop_by_day)[1] <- "date"

pop_by_day <- pop_by_day %>% gather(type, value, -date)

ggplot(pop_by_day) + geom_line(aes(x = date, y=value, col=type))

ggplot(pop_by_day)+ geom_smooth(aes(x = date, y=value, col=type))
```


#### Let's see some of the Trump's Favorite Tweets

```{r}
Trumptweets[order(Trumptweets$favorite_count, decreasing = T), ][1:5, c("favorite_count", "text")]
```

#### Let's see some of the Trump's Retweets Tweets

```{r}
Trumptweets[order(Trumptweets$retweet_count, decreasing = T), ][1:5, c("retweet_count", "text")]
```

#### Now lets see Obama's Tweet Popularity 

```{r}
pop_by_day <- aggregate(cbind(favorite_count, retweet_count) ~ as.Date(date), data = Obamatweets, FUN = median)
head(pop_by_day)

colnames(pop_by_day)[1] <- "date"

pop_by_day <- pop_by_day %>% gather(type, value, -date)

ggplot(pop_by_day) + geom_line(aes(x = date, y=value, col=type))

ggplot(pop_by_day)+ geom_smooth(aes(x = date, y=value, col=type))

#### Popular tweets
#####Favorites
Obamatweets[order(Obamatweets$favorite_count, decreasing = T), ][1:5, c("favorite_count", "text")]

#####Retweets
Obamatweets[order(Obamatweets$retweet_count, decreasing = T), ][1:5, c("retweet_count", "text")]
Obamatweets[c(680,768,868),]

```

#### Who did Trump Tweet at? 

```{r}
trumptweets_at <- regmatches(Trumptweets$text, gregexec("@[^[:space:]]+" , Trumptweets$text))
trumptweets_at <- unlist(trumptweets_at)
trumptweets_at
```

#### Plot of Trumps top 10 tweets at 

```{r}
Trumptweets_at_tab <- sort(table(trumptweets_at), decreasing = T)
barplot(Trumptweets_at_tab[1:10], horiz = TRUE, main = "Trump Top Tweets @")
```
#### Let's make th egraph more readable 

```{r}
par(las = 1, mar = c(5.1,8,4.1,2.1))
barplot(Trumptweets_at_tab[15:1], horiz = TRUE,col = "blue", las = 1, cex.names = .8, main = "Trump Top Tweets @")
```

#### let's do some more data cleaning and processing to remove unwanted characters from the @ names

```{r}
trumptweets_at2 <- gsub("Trump*." ,"Trump", trumptweets_at)
trumptweets_at2 <- gsub("Trump*." ,"Trump", trumptweets_at2)
trumptweets_at2 <- gsub("[.]", "", trumptweets_at2)
trumptweets_at2 <- gsub("[:]", "", trumptweets_at2)
```

```{r}
trumptweets_at2_tab2 <- sort(table(trumptweets_at2), decreasing = T)
par(las = 1, mar = c(5.1,8,4.1,2.1))
barplot(trumptweets_at2_tab2[15:1], horiz = TRUE,col = "blue", las = 1, cex.names = .8, main = "Trump Top Tweets @")
```

#### Who did Obama Tweet at? 

```{r}
regmatches(Obamatweets$text, gregexpr("@[^[:space:]]+", Obamatweets$text))

Obamatweets_at <- regmatches(Obamatweets$text, gregexpr("@[^[:space:]]+", Obamatweets$text))

Obamatweets_at <- unlist(Obamatweets_at)
Obamatweets_at

Obamatweets_at_tab <- sort(table(Obamatweets_at), decreasing = TRUE)

par(las = 1, mar = c(5.1,8,4.1,2.1))
barplot(Obamatweets_at_tab[15:1], horiz = TRUE, col = 'blue', las = 1, cex.names = .8, main = "Obama Top Tweets @")

#Again, remove : and . and ,
Obamatweets_at2 <- gsub("[.]|[,]|[:]", "", Obamatweets_at)

Obamatweets_at_tab2 <- sort(table(Obamatweets_at2), decreasing = TRUE)
barplot(Obamatweets_at_tab2[15:1], horiz = TRUE, col = 'blue', las = 1, cex.names =.8, main = "Obama Top Tweets @")
```


#### Tweets To Words 

##### Let's look at Trump's Tweets 

##### First We will do some cleaning like removing links and ampersands

```{r}
Trumptweets$text_clean <- trimws(gsub("@[^[:space:]]+", "", Trumptweets$text))
Trumptweets$text_clean <- gsub("https://t.co/[A-Za-z0-9]+", "", Trumptweets$text_clean)
Trumptweets$text_clean <- gsub("&amp;", "&", Trumptweets$text_clean, fixed = TRUE)
Trumptweets$text_clean <- gsub("창|탑", "", Trumptweets$text_clean)

tweet_words <- Trumptweets
tweet_words <- unnest_tokens(tweet_words, word, text_clean, strip_punct = TRUE)
tweet_words <- tweet_words[grep("[a-z]", tweet_words$word),]
tweet_words <- tweet_words[!(tweet_words$word %in% stop_words$word),]
```

#### which words?

```{r}
head(sort(tweet_words$word), 100)
```


```{r}
tail(sort(tweet_words$word), 100)
```

#### Comparing word frequency 

```{r}
word_freq <- sort(table(tweet_words$word), decreasing = T)
word_freq[1:50]
```

#### Let's make a plot of the word frequency

```{r}
par(las =1)
par(mar = c(5,8,4,2), cex = .8)

barplot(word_freq[15:1], horiz = T, col = "red", main = "Top 15 Tweeted words", xlab = "Count")
```

#### Let's make a word cloud to present the tweeted words frequency 

```{r}
wordcloud(names(word_freq), word_freq, min.freq = 40, random.order = F, scale = c(3,0.5), colors = brewer.pal(8, "Dark2"))
```

##### Let's look at Obama's Tweets 

##### First We will do some cleaning like removing links and ampersands

```{r}

# remove links and clean ampersands
Obamatweets$text_clean <- trimws(gsub("@[^[:space:]]+", "", Obamatweets$text))
Obamatweets$text_clean <- gsub("https://t.co/[A-Za-z0-9]+", "", Obamatweets$text_clean)
Obamatweets$text_clean <- gsub("&amp;", "&", Obamatweets$text_clean, fixed = TRUE)
Obamatweets$text_clean <- gsub("창|탑", "", Obamatweets$text_clean)

tweet_words <- Obamatweets
tweet_words <- unnest_tokens(tweet_words, word, text_clean, strip_punct = TRUE)
tweet_words <- tweet_words[grep("[a-z]", tweet_words$word),]
tweet_words <- tweet_words[!(tweet_words$word %in% stop_words$word),]

# which words?
head(sort(tweet_words$word), 100)
tail(sort(tweet_words$word), 100)
```


Comparing word frequencies

```{r}
word_freq <- sort(table(tweet_words$word), decreasing = TRUE)
word_freq[1:50]
```
Wordcloud

```{r}
#decrease min.freq to increase included words
wordcloud(names(word_freq), word_freq, min.freq = 30, random.order = FALSE,
          scale = c(3, 0.5), colors = brewer.pal(8, "Dark2"))
```

### Sentiment Analysis 

#### Let's get the sentiment of Trump's tweet 

```{r}
sent_df <- get_nrc_sentiment(Trumptweets$text_clean)
head(sent_df)

```


```{r}
sent_value <- get_sentiment(Trumptweets$text_clean)
Trumptweets$sent <- sent_value
```

#### What's the sentiment score for his top 15 tweeted words?

```{r}
sent_value <- get_sentiment(names(word_freq[20:1]))
names(sent_value) <- names(word_freq[20:1])
par(mar = c(1,7,2,1))
barplot(sent_value, horiz = TRUE, las = 1)
```
Hmm, maybe we should remove all the zeros

```{r}
#maybe should get rid of zeros
sent_value <- get_sentiment(names(word_freq))
names(sent_value) <- names(word_freq)
sent_value <- sent_value[sent_value != 0]
par(mar = c(1, 7, 2, 1))
barplot(sent_value[20:1], horiz = TRUE, col = 'red', las = 1)

```
#### Let's get rid of retweets since these aren't his words.

```{r}
Trumptweets_words <- Trumptweets[Trumptweets$is_retweet == "FALSE", ]
```

#### Most negative tweets:

```{r}
worst5 <- order(Trumptweets_words$sent)[1:5]
Trumptweets_words$text_clean[worst5]
```

#### Most positive tweets:

```{r}
best5 <- order(Trumptweets_words$sent, decreasing = TRUE)[1:5]
Trumptweets_words$text_clean[best5]
```

#### A quick look at sentiment over time:

```{r}
Trumptweets_words$date2 <- date(Trumptweets_words$date)
ggplot(Trumptweets_words) + geom_smooth(aes(x = date2, y = sent))
```


#### Now, Lets have a look at sentiment analysis of Obama's Tweets 

```{r}

#Let's get the sentiments of current tweets - this takes a second to run
sent_df <- get_nrc_sentiment(Obamatweets$text_clean)
head(sent_df)
sent_value <- get_sentiment(Obamatweets$text_clean)
Obamatweets$sent <- sent_value
```

What's the sentiment score for his top 15 tweeted words without the zeros?

```{r}
sent_value <- get_sentiment(names(word_freq))
names(sent_value) <- names(word_freq)
sent_value <- sent_value[sent_value != 0]
par(mar = c(1, 7, 2, 1))
barplot(sent_value[20:1], horiz = TRUE, col = 'red', las = 1)
```

Let's get rid of retweets since these aren't his words.

```{r}
Obamatweets_words <- Obamatweets[Obamatweets$is_retweet == FALSE, ]
```

Most negative tweets:

```{r}
worst5 <- order(Obamatweets_words$sent)[1:5]
Obamatweets_words$text_clean[worst5]
```

Most positive tweets:

```{r}
best5 <- order(Obamatweets_words$sent, decreasing = TRUE)[1:5]
Obamatweets_words$text_clean[best5]
```
A quick look at sentiment over time:

```{r}
Obamatweets_words$date2 <- date(Obamatweets_words$date)
ggplot(Obamatweets_words) + geom_smooth(aes(x = date2, y = sent))
```


The end!
